### 2023-07-05 chapters/background.tex
addressing comments

\subsection{Rule-based Classification}
A classifier $ \Rule: \mathbf{X} \rightarrow \widehat{Y} \in  \{0,1\} $ is a  function that takes a feature vector as input and outputs the predicted class label $\widehat{Y}$.  %We define two rules  $ \R_1 $ and $ \R_2 $ to be  equivalent  if $ \forall i,\R_1(\X_i)=\R_2(\X_i) $. 
In a rule-based classifier such as a CNF formula, we view $ \mathcal{R} $ as a propositional formula defined over Boolean features $ \mathbf{X} $ such that, if $ \Rule $ becomes true for an input, the prediction $ \widehat{Y} = 1 $ and $ \widehat{Y} = 0 $, otherwise.  The goal is to design $\Rule$ not only to approximate the training data, but also to generalize to unseen samples arising from the same distribution. Additionally, we prefer to learn a \emph{sparse} $ \Rule $ in order to favor interpretability. We define the structural complexity of $ \Rule $, referred to as \emph{rule-size}, in terms of the number of literals it contains. Let $\clause(\Rule,i)$ denote  the  $i^\text{th}$ clause of $\Rule$ and $ |\clause(\Rule,i)| $ be  the number of literals in $\clause(\Rule,i)$. Thus, the size of the classifier is $|\Rule| =  \Sigma_i |\clause(\Rule,i)| $, which is the sum of literals in all clauses in $ \Rule $.

@@

\subsection{Fairness Metrics}

A classifier $\alg$ that solely optimizes accuracy, i.e. the average number of times $\widehat{Y} = Y$, may discriminate certain compound sensitive groups over others~\cite{chouldechova2020snapshot}. In the following, we describe two well-known fairness definitions: group fairness and causal fairness. To this end, we denote $ f(\alg, \mathcal{D}) $ to quantify the fairness of the classifier $ \alg $ given the distribution of features $ \mathcal{D} $. Alternatively, a fairness metric can be computed on a finite dataset instead of on the distribution. In that case, we use the notation $ f(\alg, \dataset) $, which is the bias of the classifier $ \alg $ on a dataset $ \dataset $.

@@

\subsection{Bayesian Network}
\label{chapter_fairness_preliminaries_BN}
In general, a Probabilistic Graphical Model~\cite{koller2009probabilistic}, and specifically a \textit{Bayesian network}~\cite{pearl1985bayesian,chavira2008probabilistic}, encodes the dependencies and conditional independence between a set of random variables. In fairness verification with correlated features, we leverage an access to a Bayesian network on sensitive and non-sensitive features $ \nonsensitive \cup \sensitive $ that represents the joint distribution on them. A Bayesian network is denoted by a pair $ (\graph, \theta)$, where $ \graph \triangleq (\graphvertices, \graphedges) $ is a DAG (Directed Acyclic Graph), and $\theta$ is a set of parameters encoding the conditional probabilities induced by the joint distribution under investigation. Each vertex $\graphvertex_i \in \graphvertices$ corresponds to a random variable. Edges $ \graphedges \in \graphvertices \times \graphvertices $ imply conditional dependencies among variables. For each variable $ \graphvertex_i \in \graphvertices $, let $ \parent(\graphvertex_i) \subseteq \graphvertices \setminus \{\graphvertex_i\} $ denote the set of parents of $ \graphvertex_i $. Given $\parent(\graphvertex_i)$ and parameters $\theta$, $ \graphvertex_i $ is independent of its other non-descendant variables in $\graph$. Thus, for the assignment $ v_i $ of $ \graphvertex_i $ and $ \mathbf{u} $ of $ \parent(\graphvertex_i) $, the aforementioned semantics of a Bayesian network encodes the joint distribution of $\graphvertices$ as:

@@

\subsection{Global Sensitivity Analysis (GSA): Variance Decomposition}
Global sensitivity analysis is an active field of research that studies how the global uncertainty in the output of a function can be attributed to the different sources of uncertainties in the input variables~\cite{saltelli2008global}.
Sensitivity analysis is an essential component for quality assurance and impact assessment of models in EU~\cite{eu}, USA~\cite{usepa}, and research communities~\cite{saltelli2020five}.
%\paragraph{Sensitivity Analysis:}
\emph{Variance-based sensitivity analysis} is a form of global sensitivity analysis, where variance is considered as the measure of uncertainty~\cite{sobol1990sensitivity,sobol2001global}. To illustrate, let us consider a real-valued function $  \fiffunc(\feature) $, where $ \feature $ is a vector of $ \numfeatures $ input variables $ \{Z_1, \dots, Z_\numfeatures\} $.
% and $ Y $ is a univariate output. 
%\blue{In sensitivity analysis, assumptions are made on the independence and uniform distribution of $ X_i $'s within the unit hypercube}, that is, $ X_i \in [0,1] $ for $ i = \{ 1, \dots, k\} $.\footnote{In our experiments, we have not considered this yet.} Let $ [\numnonsensitive] \triangleq \{1,2,\dots, \numnonsensitive\} $. Then, 
Now, we decompose $ \fiffunc(\feature) $ among the subsets of inputs, such that:
\begin{align}
	\fiffunc(\feature) &= \fiffunc_0 + \sum_{i=1}^{\numfeatures} \fiffunc_{\{i\}}(Z_i) +  \sum_{i < j}^{\numfeatures} \fiffunc_{\{i,j\}}(Z_i, Z_j)  + \cdots\notag + \fiffunc_{\{1, 2, \dots, \numfeatures\}} (Z_1, Z_2, \dots, Z_{\numfeatures})\\
	&= \fiffunc_0 +  \sum_{\mathbf{S} \subseteq [\numfeatures]} \fiffunc_{\mathbf{S}}(\feature_{\mathbf{S}})\label{fairness_fairXplainer_eq:functional_decomposition_set_notation}
\end{align}
%	$S_> = \{{S_0}, \ldots, {S_{|s|}}\}$ i.e. a monotonically increasing ordering of $|S|$ features for a given starting feature $Z_{S_0}$ such that $S_0 <S_1 <\ldots <S_{|S|}$.
The standard condition of this decomposition is the orthogonality of each term in the right-hand side of Eq.~\eqref{fairness_fairXplainer_eq:functional_decomposition_set_notation}~\cite{sobol1990sensitivity}. In this decomposition, $ \fiffunc_0 $ is a constant, $ \fiffunc_{\{i\}} $ is a function of $ Z_i $, $ \fiffunc_{\{i,j\}} $ is a function of $ Z_i $ and $ Z_j $, and so on. Adhering to the set-based notations, we denote by $ g_{\mathbf{S}} $ a function of a non-empty subset of variables $ \feature_{\mathbf{S}}  \triangleq \{Z_i \mid i \in \mathbf{S}\} \subseteq \feature $, where $ \mathbf{S} = \{S_i \mid 1 \le |S_i| \le \numfeatures\} \subseteq [\numfeatures] $ is a non-empty subset of indices with $ [\numfeatures] \triangleq \{1,2,\dots, \numfeatures\}  $.  Here, $|\mathbf{S}|=1$ quantifies an individual variable's effect while $|\mathbf{S}|>1$ quantifies the higher-order intersectional effect of variables.

>>>>

\subsection{Rule-based Classification}
\label{chapter_interpretability_preliminaries_rule_based_classification}
A classifier $ \Rule: \mathbf{X} \rightarrow \widehat{Y} \in  \{0,1\} $ is a  function that takes a feature vector as input and outputs the predicted class label $\widehat{Y}$.  %We define two rules  $ \R_1 $ and $ \R_2 $ to be  equivalent  if $ \forall i,\R_1(\X_i)=\R_2(\X_i) $. 
In a rule-based classifier such as a CNF formula, we view $ \mathcal{R} $ as a propositional formula defined over Boolean features $ \mathbf{X} $ such that, if $ \Rule $ becomes true for an input, the prediction $ \widehat{Y} = 1 $ and $ \widehat{Y} = 0 $, otherwise.  The goal is to design $\Rule$ not only to approximate the training data, but also to generalize to unseen samples arising from the same distribution. Additionally, we prefer to learn a \emph{sparse} $ \Rule $ in order to favor interpretability. We define the structural complexity of $ \Rule $, referred to as \emph{rule size}, in terms of the number of literals it contains. Let $\clause(\Rule,i)$ denote  the  $i^\text{th}$ clause of $\Rule$ and $ |\clause(\Rule,i)| $ be  the number of literals in $\clause(\Rule,i)$. Thus, the size of the classifier is $|\Rule| =  \Sigma_i |\clause(\Rule,i)| $, which is the sum of literals in all clauses in $ \Rule $.

@@

\subsection{Fairness Metrics}
\label{chapter_fairness_preliminaries_fairness_metrics}
A classifier $\alg$ that solely optimizes accuracy, i.e. the average number of times $\widehat{Y} = Y$, may discriminate certain compound sensitive groups over others~\cite{chouldechova2020snapshot}. In the following, we describe two well-known fairness definitions: group fairness and causal fairness. To this end, we denote $ f(\alg, \mathcal{D}) $ to quantify the fairness of the classifier $ \alg $ given the distribution of features $ \mathcal{D} $. Alternatively, a fairness metric can be computed on a finite dataset instead of on the distribution. In that case, we use the notation $ f(\alg, \dataset) $, which is the bias of the classifier $ \alg $ on a dataset $ \dataset $.

@@

\section{Bayesian Network}
\label{chapter_fairness_preliminaries_BN}
In general, a Probabilistic Graphical Model~\cite{koller2009probabilistic}, and specifically a \textit{Bayesian network}~\cite{pearl1985bayesian,chavira2008probabilistic}, encodes the dependencies and conditional independence between a set of random variables. In fairness verification with correlated features, we leverage an access to a Bayesian network on sensitive and non-sensitive features $ \nonsensitive \cup \sensitive $ that represents the joint distribution on them. A Bayesian network is denoted by a pair $ (\graph, \theta)$, where $ \graph \triangleq (\graphvertices, \graphedges) $ is a DAG (Directed Acyclic Graph), and $\theta$ is a set of parameters encoding the conditional probabilities induced by the joint distribution under investigation. Each vertex $\graphvertex_i \in \graphvertices$ corresponds to a random variable. Edges $ \graphedges \in \graphvertices \times \graphvertices $ imply conditional dependencies among variables. For each variable $ \graphvertex_i \in \graphvertices $, let $ \parent(\graphvertex_i) \subseteq \graphvertices \setminus \{\graphvertex_i\} $ denote the set of parents of $ \graphvertex_i $. Given $\parent(\graphvertex_i)$ and parameters $\theta$, $ \graphvertex_i $ is independent of its other non-descendant variables in $\graph$. Thus, for the assignment $ v_i $ of $ \graphvertex_i $ and $ \mathbf{u} $ of $ \parent(\graphvertex_i) $, the aforementioned semantics of a Bayesian network encodes the joint distribution of $\graphvertices$ as:

@@

\section{Global Sensitivity Analysis (GSA): Variance Decomposition}
Global sensitivity analysis is an active field of research that studies how the global uncertainty in the output of a function can be attributed to the different sources of uncertainties in the input variables~\cite{saltelli2008global}.
Sensitivity analysis is an essential component for quality assurance and impact assessment of models in EU~\cite{eu}, USA~\cite{usepa}, and research communities~\cite{saltelli2020five}.
%\paragraph{Sensitivity Analysis:}
\emph{Variance-based sensitivity analysis} is a form of global sensitivity analysis, where variance is considered as the measure of uncertainty~\cite{sobol1990sensitivity,sobol2001global}. To illustrate, let us consider a real-valued function $  \fiffunc(\feature) $, where $ \feature $ is a vector of $ \numfeatures $ input variables $ \{Z_1, \dots, Z_\numfeatures\} $.
% and $ Y $ is a univariate output. 
%\blue{In sensitivity analysis, assumptions are made on the independence and uniform distribution of $ X_i $'s within the unit hypercube}, that is, $ X_i \in [0,1] $ for $ i = \{ 1, \dots, k\} $.\footnote{In our experiments, we have not considered this yet.} Let $ [\numnonsensitive] \triangleq \{1,2,\dots, \numnonsensitive\} $. Then, 
Now, we decompose $ \fiffunc(\feature) $ among the subsets of inputs, such that:
\begin{align}
	\fiffunc(\feature) &= \fiffunc_0 + \sum_{i=1}^{\numfeatures} \fiffunc_{\{i\}}(Z_i) +  \sum_{i < j}^{\numfeatures} \fiffunc_{\{i,j\}}(Z_i, Z_j)  + \cdots\notag + \fiffunc_{\{1, 2, \dots, \numfeatures\}} (Z_1, Z_2, \dots, Z_{\numfeatures})\\
	&= \fiffunc_0 +  \sum_{\mathbf{S} \subseteq [\numfeatures]} \fiffunc_{\mathbf{S}}(\feature_{\mathbf{S}})\label{fairness_fairXplainer_eq:functional_decomposition_set_notation}
\end{align}
%	$S_> = \{{S_0}, \ldots, {S_{|s|}}\}$ i.e. a monotonically increasing ordering of $|S|$ features for a given starting feature $Z_{S_0}$ such that $S_0 <S_1 <\ldots <S_{|S|}$.
The standard condition of this decomposition is the orthogonality of each term in the right-hand side of Eq.~\eqref{fairness_fairXplainer_eq:functional_decomposition_set_notation}~\cite{sobol1990sensitivity}. In this decomposition, $ \fiffunc_0 $ is a constant, $ \fiffunc_{\{i\}} $ is a function of $ Z_i $, $ \fiffunc_{\{i,j\}} $ is a function of $ Z_i $ and $ Z_j $, and so on. Adhering to the set-based notations, we denote by $ g_{\mathbf{S}} $ a function of a non-empty subset of variables $ \feature_{\mathbf{S}}  \triangleq \{Z_i \mid i \in \mathbf{S}\} \subseteq \feature $, where $ \mathbf{S} = \{S_i \mid 1 \le |S_i| \le \numfeatures\} \subseteq [\numfeatures] $ is a non-empty subset of indices with $ [\numfeatures] \triangleq \{1,2,\dots, \numfeatures\}  $.  Here, $|\mathbf{S}|=1$ quantifies an individual variable's effect while $|\mathbf{S}|>1$ quantifies the higher-order intersectional effect of variables.

### 2023-07-05 chapters/conclusion.tex
addressing comments

To demonstrate the efficacy of our methods, we have constructed open-source tools and conducted experiments on real-world datasets in machine learning. In the context of interpretable rule-based machine learning, we have developed an incremental learning framework, known as {\imli}, that scales classification to million-size datasets while maintaining competitive prediction accuracy and rule-size compared to existing rule-based classifiers. Additionally, in our pursuit of more expressive yet interpretable classifiers, we have introduced another learning framework, called {\crr}, for logical relaxed classification rules that are based on incremental learning. Experimental results show that {\crr} is capable of learning more concise and accurate rule-based classifiers.

>>>>

To demonstrate the efficacy of our methods, we have constructed open-source tools and conducted experiments on real-world datasets in machine learning. In the context of interpretable rule-based machine learning, we have developed an incremental learning framework, known as {\imli}, that scales classification to million-size datasets while maintaining competitive prediction accuracy and rule size compared to existing rule-based classifiers. Additionally, in our pursuit of more expressive yet interpretable classifiers, we have introduced another learning framework, called {\crr}, for logical relaxed classification rules that are based on incremental learning. Experimental results show that {\crr} is capable of learning more concise and accurate rule-based classifiers.

### 2023-07-05 chapters/fairness/fif/experiments.tex
addressing comments

\paragraph{Execution Time: {\fairXplainer} vs.\ SHAP} We compare the execution time of {\fairXplainer} vs.\ SHAP in a cactus plot in Figure~\ref{fairness_fairXplainer_fig:execution_time_cactus_plot}, where a point $ (x, y) $ denotes that a method computes FIFs of $ y $ many fairness instances within $ x $ seconds. We consider $ 480 $ fairness instances from $ 4 $ datasets constituting $ 24 $ distinct combinations of sensitive features, $ 4 $ classifiers, and $ 5 $ cross-validation folds. In Figure~\ref{fairness_fairXplainer_fig:execution_time_cactus_plot},  {\fairXplainer} with $ \lambda = 1 $ is faster than  $ \lambda = 2 $. For example, within $ 10 $ seconds, {\fairXplainer} with $ \lambda = 1 $ solves $ 443 $ instances vs.\ $ 41 $ instances with $ \lambda = 2 $.  In addition, {\fairXplainer} with $ \lambda = 1 $ solves all $ 480 $ instances compared to $ 360 $ instances solved by SHAP. Thus, {\fairXplainer} with $ \lambda = 1 $ demonstrates its higher efficiency in estimating individual FIFs compared to SHAP. While estimating intersectional FIFs, {\fairXplainer} also demonstrates its practical applicability by solving $ 350 $ instances within  $ 160 $ seconds. \textit{Therefore, {\fairXplainer} demonstrates computational efficiency in interpreting group fairness metrics of real-world datasets and classifiers.}

>>>>

\paragraph{Execution Time: {\fairXplainer} vs.\ SHAP} We compare the execution time of {\fairXplainer} vs.\ SHAP in a cactus plot in Figure~\ref{fairness_fairXplainer_fig:execution_time_cactus_plot}, where a point $ (x, y) $ denotes that a method computes FIFs of $ x $ many fairness instances within $ y $ seconds. We consider $ 480 $ fairness instances from $ 4 $ datasets constituting $ 24 $ distinct combinations of sensitive features, $ 4 $ classifiers, and $ 5 $ cross-validation folds. In Figure~\ref{fairness_fairXplainer_fig:execution_time_cactus_plot},  {\fairXplainer} with $ \lambda = 1 $ is faster than  $ \lambda = 2 $. For example, within $ 10 $ seconds, {\fairXplainer} with $ \lambda = 1 $ solves $ 443 $ instances vs.\ $ 41 $ instances with $ \lambda = 2 $.  In addition, {\fairXplainer} with $ \lambda = 1 $ solves all $ 480 $ instances compared to $ 360 $ instances solved by SHAP. Thus, {\fairXplainer} with $ \lambda = 1 $ demonstrates its higher efficiency in estimating individual FIFs compared to SHAP. While estimating intersectional FIFs, {\fairXplainer} also demonstrates its practical applicability by solving $ 350 $ instances within  $ 160 $ seconds. \textit{Therefore, {\fairXplainer} demonstrates computational efficiency in interpreting group fairness metrics of real-world datasets and classifiers.}

### 2023-07-05 chapters/fairness/fif/framework.tex
addressing comments

\subsection{Fairness Metrics as the Variance of Prediction}
\label{fairness_fairXplainer_sec:metric_as_variance}

First, we observe that the random variable central to computing statistical parity is $ \mathds{1}[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}] $.
We refer to this indicator function as \textit{Conditional Positive Prediction (CPP)} of a classifier. Now, we express statistical parity as a functional of the probability of CPPs for different sensitive groups~\cite{benesse2021fairness}. For brevity, we defer similar formulations for the other fairness metrics, i.e. equalized odds and predictive parity, to Appendix~\ref{fairness_fairXplainer_app:pp}. %CPP has an additional conditional $ Y = y $.}
%For indicator functions, probability and expectation are equal.

@@

>>>>

\subsection{Fairness Metrics as the Variance of Prediction}
\label{fairness_fairXplainer_sec:metric_as_variance}
First we recall the definition of statistical parity in Chapter~\ref{chapter_fairness_preliminaries_fairness_metrics}.
\[ f_{\mathsf{SP}}(\alg, \mathbf{D}) \triangleq \max_{\mathbf{a}}\Pr[\widehat{Y} =1 | \mathbf{A} = \mathbf{a}] - \min_{\mathbf{a}}\Pr[\widehat{Y} =1 | \mathbf{A} = \mathbf{a}] \]

Therefore, the random variable central to computing statistical parity is $ \mathds{1}[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}] $.
We refer to this indicator function as \textit{Conditional Positive Prediction (CPP)} of a classifier. Now, we express statistical parity as a functional of the probability of CPPs for different sensitive groups~\cite{benesse2021fairness}. For brevity, we defer similar formulations for the other fairness metrics, i.e. equalized odds and predictive parity, to Appendix~\ref{fairness_fairXplainer_app:pp}. %CPP has an additional conditional $ Y = y $.}
%For indicator functions, probability and expectation are equal.

@@

\begin{remark}
	For detecting degenerate cases, we require to compute $ \Pr[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}_{\max} ]  $ and $ \Pr[\widehat{Y} = 1 \mid \sensitive = \mathbf{a}_{\min} ]  $ on the dataset $ \mathbf{D} $. Hence, the detection can be performed in a straightforward manner.
\end{remark}

### 2023-07-05 chapters/interpretability/imli/appendix.tex
addressing comments

In Figure~\ref{fig:interpretability_imli_additional_fig:all_preformances_non_incremental_vs_incremental}, we compare both non-incremental and incremental encoding in terms of median training time, rule-size, test accuracy, and train accuracy over ten cross validation folds. In all datasets, the incremental encoding demonstrates higher efficiency than the non-incremental encoding by requiring less training time. In large datasets, such as Magic, Credit, and Adult, the non-incremental encoding always times out, thereby producing zero-size rules because of failing to solve the MaxSAT query. As a result, the accuracy of the non-incremental encoding is also less than the incremental encoding in large datasets. In small datasets, the incremental encoding demonstrates a competitive accuracy and rule-size with the non-incremental encoding. Therefore, the incremental encoding in {\imli} demonstrates its effectiveness in training time, accuracy, and rule-size over the non-incremental encoding.

@@

	\caption[Performance comparison between non-incremental vs.\ incremental encoding]{Comparison of training time, rule-size, and test and train accuracy between non-incremental vs.\ incremental encoding in {\imli}. In $ X $-axis, the dimension of the dataset (\#samples $ \mid $  \#features) is shown below dataset name. In large datasets, non-incremental encoding often times out and learns zero-size classification rules yielding less test and train accuracy than the incremental encoding.}
	\label{fig:interpretability_imli_additional_fig:all_preformances_non_incremental_vs_incremental}
\end{figure}

>>>>

In Figure~\ref{fig:interpretability_imli_additional_fig:all_preformances_non_incremental_vs_incremental}, we compare both non-incremental and incremental encoding in terms of median training time, rule size, test accuracy, and train accuracy over ten cross validation folds. In all datasets, the incremental encoding demonstrates higher efficiency than the non-incremental encoding by requiring less training time. In large datasets, such as Magic, Credit, and Adult, the non-incremental encoding always times out, thereby producing zero-size rules because of failing to solve the MaxSAT query. As a result, the accuracy of the non-incremental encoding is also less than the incremental encoding in large datasets. In small datasets, the incremental encoding demonstrates a competitive accuracy and rule size with the non-incremental encoding. Therefore, the incremental encoding in {\imli} demonstrates its effectiveness in training time, accuracy, and rule size over the non-incremental encoding.

@@

	\caption[Performance comparison between non-incremental vs.\ incremental encoding]{Comparison of training time, rule size, and test and train accuracy between non-incremental vs.\ incremental encoding in {\imli}. In $ X $-axis, the dimension of the dataset (\#samples $ \mid $  \#features) is shown below dataset name. In large datasets, non-incremental encoding often times out and learns zero-size classification rules yielding less test and train accuracy than the incremental encoding.}
	\label{fig:interpretability_imli_additional_fig:all_preformances_non_incremental_vs_incremental}
\end{figure}

### 2023-07-05 chapters/interpretability/imli/baseline.tex
addressing comments

We consider two types of Boolean variables: (i) \emph{feature} variables $ \bool $ corresponding to input features and (ii) \emph{error} variables $ \noise $ corresponding to the classification error of samples. We define a Boolean variable $ \bool^i_j $ that becomes true if feature $ X_j $ appears in the $ i^\text{th} $ clause of $ \Rule $, thereby contributing to an increase in the rule-size of $ \Rule $, and $ \bool^i_j $ is assigned false, otherwise. Moreover, we define an error variable $ \noise_l $ to attribute to whether the $ l^\text{th} $ sample $ (\mathbf{x}^{(l)}, y^{(l)}) $ is classified correctly or not. Specifically, $ \noise_l $ becomes true if $ (\mathbf{x}^{(l)}, y^{(l)}) $ is misclassified, and becomes false otherwise.  We next discuss the MaxSAT encoding to solve the classification problem.

@@

In this example, we consider regularizer $ \lambda = 0.1 $, thereby setting the weight on accuracy as $ \clauseweight{E_i} = 1 $ and rule-sparsity weight as $ \clauseweight{S_j^i} = 0.1 $. Intuitively, the penalty for misclassifying a sample is $ 10 $ times than the penalty for adding a feature in the classifier. For this MaxSAT query, the optimal solution classifies all samples correctly by assigning four feature variables $ \{\bool^1_1, \bool^1_4, \bool^2_2, \bool^2_3\} $ to true. Therefore, by applying Construction~\ref{interpretability_imli_construction:rule}, the classifier is $ (X_1 \vee \neg X_2)  \wedge (\neg X_1 \vee X_2) $.\footnote{The presented MaxSAT-based formulation does not learn a CNF classifier with both $ X_i $ and $ \neg X_i $ in the same clause. The reason is that a clause with both $ X_i $ and $ \neg X_i $ connected by OR $ (\vee) $ does not increase accuracy but increases rule-size and hence, this is not an optimal classifier.}

>>>>

We consider two types of Boolean variables: (i) \emph{feature} variables $ \bool $ corresponding to input features and (ii) \emph{error} variables $ \noise $ corresponding to the classification error of samples. We define a Boolean variable $ \bool^i_j $ that becomes true if feature $ X_j $ appears in the $ i^\text{th} $ clause of $ \Rule $, thereby contributing to an increase in the rule size of $ \Rule $, and $ \bool^i_j $ is assigned false, otherwise. Moreover, we define an error variable $ \noise_l $ to attribute to whether the $ l^\text{th} $ sample $ (\mathbf{x}^{(l)}, y^{(l)}) $ is classified correctly or not. Specifically, $ \noise_l $ becomes true if $ (\mathbf{x}^{(l)}, y^{(l)}) $ is misclassified, and becomes false otherwise.  We next discuss the MaxSAT encoding to solve the classification problem.

@@

In this example, we consider regularizer $ \lambda = 0.1 $, thereby setting the weight on accuracy as $ \clauseweight{E_i} = 1 $ and rule-sparsity weight as $ \clauseweight{S_j^i} = 0.1 $. Intuitively, the penalty for misclassifying a sample is $ 10 $ times than the penalty for adding a feature in the classifier. For this MaxSAT query, the optimal solution classifies all samples correctly by assigning four feature variables $ \{\bool^1_1, \bool^1_4, \bool^2_2, \bool^2_3\} $ to true. Therefore, by applying Construction~\ref{interpretability_imli_construction:rule}, the classifier is $ (X_1 \vee \neg X_2)  \wedge (\neg X_1 \vee X_2) $.\footnote{The presented MaxSAT-based formulation does not learn a CNF classifier with both $ X_i $ and $ \neg X_i $ in the same clause. The reason is that a clause with both $ X_i $ and $ \neg X_i $ connected by OR $ (\vee) $ does not increase accuracy but increases rule size and hence, this is not an optimal classifier.}
